Reference label transfer exploration for `r params$sample_id`

### 2024-08-13 analyses/cell-type-wilms-tumor-06/notebook_template/01_fetal_reference_Stewart_Seuratv4.Rmd
add to PR #706

---
title: "Reference label transfer exploration for `r params$sample_id`"
author: "Maud PLASCHKA"
date: '2024-08-13'
params:
  current: '2024-07-08'
  scpca_project_id: "SCPCP000006"
  sample_id: "SCPCS000176"
  ref_url: "https://datasets.cellxgene.cziscience.com/40ebb8e4-1a25-4a33-b8ff-02d1156e4e9b.rds"
  seed: 12345
output: 
  html_document: 
    toc: yes
    toc_float: yes
    code_folding: hide
    highlight: pygments
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE,
                      warnings=FALSE
                      )
```

@@

##########################################################################################################
In this notebook, we will test the human fetal kidney atlas from Stewart et al. using the Azimuth workflow. 
##########################################################################################################

@@

```{r packages, message=FALSE, warning=FALSE}
library("Seurat")
library(sctransform)
library(Azimuth)
library(SCpubr)
library(tidyverse)
library(patchwork)
library(SingleCellExperiment)

@@

```{r base paths, eval=TRUE}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)

@@

```{r path_to_reference}
path_to_reference <- file.path(module_base, "scratch", "fetal_full.rds")
```

@@

This notebook has two main outputs

#### Azimuth compatible reference

The reference data are downloaded from the `params$ref_url` and transiently saved in the `scratch`folder.
The azimuth compatible fetal kidney reference will be saved in the `results` folder from the module, `reference` subfolder.

Please note that the creation of the Azimuth reference cannot be done within RMarkdown. 
We thus set `include = FALSE` to the corresponding chunk.
Please, before running this code either :

- download the azimuth compatible reference from S3 bucket or
- run manualy chunk by chunk the report to create and save the reference once.

```{r path_to_output}
path_to_output <- file.path(module_base, "results", "references")
```

#### The html notebook saved afetr rendering this notebook (knit).
Will be saved in the `notebook/00-reference` directory.

@@

### Create a azimuth compatible reference

Note: not shown here because the `AzimuthReference` function do not allow knitr.
The reference should be build "manually" prior to knitr the report.

```{r create_ref, eval=FALSE, fig.height=7, fig.width=12, message=FALSE, warning=FALSE, include=FALSE, out.width='100%'}
download.file(params$ref_url, path_to_reference)
seurat <- readRDS(path_to_reference)

options(future.globals.maxSize= 891289600000)
s <- SCTransform(seurat, verbose = FALSE, method = "glmGamPoi", conserve.memory = TRUE)
s <- RunPCA(s, npcs = 50, verbose = FALSE)
s <- RunUMAP(s, dims = 1:50, verbose = FALSE, return.model = TRUE)

Fetal_kidney <- AzimuthReference(
  s,
  refUMAP = "umap",
  refDR = "pca",
  refAssay = "SCT",
  dims = 1:50,
  k.param = 31,
  plotref = "umap",
  plot.metadata = NULL,
  ori.index = NULL,
  colormap = NULL,
  assays = NULL,
  metadata = c("compartment", "cell_type"),
  reference.version = "0.0.0",
  verbose = FALSE
)

d2 <- do_DimPlot(Fetal_kidney, reduction = "refUMAP", dims = c(1,2), group.by = "compartment", label = TRUE, repel = TRUE) + NoLegend() + ggtitle("umap reduction after SCTransform and Seurat workflow")
d3 <- do_DimPlot(Fetal_kidney, reduction = "refUMAP", dims = c(1,2), group.by = "cell_type", label = TRUE, repel = TRUE) + NoLegend()

d2 | d3

if (file.exists(path_to_output)){
   SaveAnnoyIndex(object = Fetal_kidney[["refdr.annoy.neighbors"]], file = file.path(path_to_output, "idx.annoy"))
    saveRDS(object = Fetal_kidney, file = file.path(path_to_output, "ref.Rds"))
} else {
    SaveAnnoyIndex(object = Fetal_kidney[["refdr.annoy.neighbors"]], file = file.path(path_to_output, "idx.annoy"))
    saveRDS(object = Fetal_kidney, file = file.path(path_to_output, "ref.Rds"))

}

# finally, we can remove the fetal_full.rds file from the marker-sets directory
file.remove(path_to_reference)
```

@@

### Azimuth annotation from fetal kidney

@@

data <-  RunAzimuth(srat, path_to_output, assay = 'RNA')

@@

d1 <- DimPlot(data, reduction = "umap", dims = c(1,2), group.by = "predicted.compartment", label = TRUE, repel = TRUE) + NoLegend()
d2 <- DimPlot(data, reduction = "umap", dims = c(1,2), group.by = "predicted.cell_type", label = TRUE, repel = TRUE) + NoLegend()

f1 <- SCpubr::do_BarPlot(sample = data,
                         group.by = "predicted.compartment",
                         split.by = "seurat_clusters",
                         position = "fill",
                         font.size = 10,
                         legend.ncol = 3) +
                         ggtitle("% cells")+
                         xlab(file)

f2 <- SCpubr::do_BarPlot(sample = data,
                         group.by = "predicted.cell_type",
                         split.by = "seurat_clusters",
                         position = "fill",
                         font.size = 10,
                         legend.ncol = 3) +
                         ggtitle("% cells")+
                         xlab(file)

>>>>

---
title: "Reference label transfer exploration for `r params$sample_id`"
author: "Maud PLASCHKA"
date: '2024-08-13'
params:
  current: "2024-07-08"
  scpca_project_id: "SCPCP000006"
  sample_id: "SCPCS000176"
  ref_url: "https://datasets.cellxgene.cziscience.com/40ebb8e4-1a25-4a33-b8ff-02d1156e4e9b.rds"
  seed: 12345
output: 
  html_document: 
    toc: yes
    toc_float: yes
    code_folding: hide
    highlight: pygments
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warnings = FALSE
)
```

@@

##########################################################################################################
In this notebook, we will test the human fetal kidney atlas from Stewart et al. using the Seurat v4 workflow. 
##########################################################################################################

@@

```{r packages, echo=TRUE, message=FALSE, warning=FALSE}
library("Seurat")
library(sctransform)
library(SCpubr)
library(tidyverse)
library(patchwork)
library(SingleCellExperiment)

@@

```{r base paths, eval=TRUE, include=TRUE}
# The base path for the OpenScPCA repository, found by its (hidden) .git directory
repository_base <- rprojroot::find_root(rprojroot::is_git_root)

@@

```{r path_to_reference}
path_to_reference <- file.path(module_base, "scratch", "fetal_full.rds")

download.file(params$ref_url, path_to_reference)
```

@@

Reports will be saved in the `notebook\00-reference` directory.

@@

@@

```{r fig.height=4, fig.width=4, warnings=FALSE}

DimPlot(srat, reduction = "umap", dims = c(1,2), group.by = "seurat_clusters", label = TRUE, repel = TRUE) + NoLegend()

### Load the reference

```{r pre_process,  fig.height=7, fig.width=12, message=FALSE, warning=FALSE, out.width='100%'}
seurat <- readRDS(path_to_reference)

options(future.globals.maxSize= 891289600000)
s <- SCTransform(seurat, verbose = FALSE, method = "glmGamPoi", conserve.memory = TRUE)
s <- RunPCA(s, npcs = 50, verbose = FALSE)
s <- RunUMAP(s, dims = 1:50, verbose = FALSE, return.model = TRUE)
```

```{r create_ref, fig.height=7, fig.width=12, message=FALSE, warning=FALSE, out.width='100%'}
d2 <- do_DimPlot(s, reduction = "umap", dims = c(1,2), group.by = "compartment", label = TRUE, repel = TRUE) + NoLegend() + ggtitle("umap reduction after SCTransform and Seurat workflow")
d3 <- do_DimPlot(s, reduction = "umap", dims = c(1,2), group.by = "cell_type", label = TRUE, repel = TRUE) + NoLegend()

d2 | d3
```

### Label transfer from fetal kidney reference using Seurat v4

@@

anchors <- FindTransferAnchors(reference = s, query = srat, dims = 1:50,
    reference.reduction = "pca")
predictions <- TransferData(anchorset = anchors, refdata = s$cell_type, dims = 1:50)
srat <- AddMetaData(srat, metadata = predictions$predicted.id, col.name =  "predicted.cell_type")

predictions <- TransferData(anchorset = anchors, refdata = s$compartment, dims = 1:50)
srat <- AddMetaData(srat, metadata = predictions$predicted.id, col.name =  "predicted.compartment")
```

@@

d1 <- DimPlot(srat, reduction = "umap", dims = c(1,2), group.by = "predicted.compartment", label = TRUE, repel = TRUE) + NoLegend()
d2 <- DimPlot(srat, reduction = "umap", dims = c(1,2), group.by = "predicted.cell_type", label = TRUE, repel = TRUE) + NoLegend()

f1 <- SCpubr::do_BarPlot(sample = srat,
                         group.by = "predicted.compartment",
                         split.by = "seurat_clusters",
                         position = "fill",
                         font.size = 10,
                         legend.ncol = 3) +
                         ggtitle("% cells")+
                         xlab(file)

f2 <- SCpubr::do_BarPlot(sample = srat,
                         group.by = "predicted.cell_type",
                         split.by = "seurat_clusters",
                         position = "fill",
                         font.size = 10,
                         legend.ncol = 3) +
                         ggtitle("% cells")+
                         xlab(file)
